\pkg{cubble}: An R Package for Organizing and Wrangling Multivariate Spatio-temporal Data

### 2022-06-14 jss/cubble-jss.tex
appendix as a single document

\begin{CodeChunk}
\begin{CodeInput}
R> cb <- as_cubble(
+   list(spatial = lga, temporal = covid),
+   key = lga, index = date, coords = c(cent_long, cent_lat)
+   )
\end{CodeInput}
\begin{CodeOutput}
! Some sites in the temporal table don't have corresponding spatial information
\end{CodeOutput}
\begin{CodeOutput}
! Some sites in the spatial table don't have corresponding temporal information
\end{CodeOutput}
\begin{CodeOutput}
! Use argument `output = "unmatch"` to check on the unmatched key
\end{CodeOutput}
\end{CodeChunk}

@@

The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world. The \textbf{cubble} dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 236 Australian stations from the GHCN starting from year 1969 and provides information also on the latitude, longitude and elevation of each of the stations. This \textbf{cubble} data is already cast into a \code{cubble} data object, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates. The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales.

@@

\code{weatherdata::climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for 639 stations in Australia from 2016 to 2020. A simple \(k\)-means algorithm based on the distance matrix is used to create 20 clusters. The dataset \code{station_nested} is a nested cubble with a cluster column indicating the group to which each station belongs. More advanced clustering algorithms can be used for other applications, as long as there is a mapping from each station to a cluster.

@@

\hypertarget{appendix}{%
\section{Appendix}\label{appendix}}

{\centering \includegraphics[width=1\linewidth,height=0.4\textheight]{../figures/diagram-keynotes/diagram-keynotes.005}

\caption[An illustration of the data model under interactive graphics with cubble]{An illustration of the data model under interactive graphics with cubble. When a point on the time series is selected, the corresponding row in the long cubble will be activated. This will link to all the rows with the same id in the long cubble and the row in the nested cubble with the same id (middle). Both plots will be updated with the full line selected and the point highlighted on the map (right).}\label{fig:illu-interactive-2}
\end{figure}
\end{CodeChunk}

>>>>

\begin{CodeChunk}
\begin{CodeInput}
R> cb <- as_cubble(
+   list(spatial = lga, temporal = covid),
+   key = lga, index = date, coords = c(cent_long, cent_lat)
+   )
\end{CodeInput}
\begin{CodeOutput}
! Some sites in the temporal table don't have spatial information
\end{CodeOutput}
\begin{CodeOutput}
! Some sites in the spatial table don't have temporal information
\end{CodeOutput}
\begin{CodeOutput}
! Use argument `output = "unmatch"` to check on the unmatched key
\end{CodeOutput}
\end{CodeChunk}

@@

The Global Historical Climatology Network (GHCN) provides daily climate measures from stations across the world. The \textbf{cubble} dataset \code{weatherdata::historical_tmax} extracts the maximum temperature for 204 Australian stations from the GHCN starting from year 1969 and provides information also on the latitude, longitude and elevation of each of the stations. This \textbf{cubble} data is already cast into a \code{cubble} data object, with \code{id} as the key, \code{date} as the index, and \code{c(longitude, latitude)} as the coordinates. The goal of this example is to compare the monthly average maximum temperature between two periods, 1971-1975 and 2016-2020, for stations in Victoria and New South Wales.

@@

\code{weatherdata::climate_full}, also extracted from the GHCN, records daily precipitation and maximum/minimum temperature for 640 stations in Australia from 2016 to 2020. A simple \(k\)-means algorithm based on the distance matrix is used to create 20 clusters. The dataset \code{station_nested} is a nested cubble with a cluster column indicating the group to which each station belongs. More advanced clustering algorithms can be used for other applications, as long as there is a mapping from each station to a cluster.

@@
